\bf Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2025-05-24 manuscript/manuscript.tex
tidied table
went over 2nd paragraph in conclusion

%&latex
\documentclass[12pt]{article}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL
\usepackage{comment}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}\label{sec:intro}

@@

\begin{table}[tbp]
\centering
\caption{Empirical sizes or rejection rates
for the test that sample follows its true distribution family for
different values of AR(1) coefficient and for different
significance levels.}
\label{table:size}
\begin{tabular}{lllrrrrrrr}
  \toprule
$F(\cdot; \theta)$ & $n$ & $\alpha$ & $\tau = -0.75$ & $\tau = -0.5$ & $\tau = -0.25$
& $\tau = 0$ & $\tau = 0.25$ & $\tau = 0.5$ & $\tau = 0.75$ \\
  \midrule
{$N(8,8)$} & {100} & 0.01 & 0.0521 & 0.0124 & 0.0112 & 0.0106 & 0.0114 & 0.0158 & 0.0563 \\
& & 0.05 & 0.1470 & 0.0628 & 0.0545 & 0.0548 & 0.0575 & 0.0725 & 0.1705 \\
& & 0.10 & 0.2277 & 0.1186 & 0.1147 & 0.1072 & 0.1154 & 0.1424 & 0.2714 \\
& {200} & 0.01 & 0.0439 & 0.0113 & 0.0090 & 0.0114 & 0.0109 & 0.0138 & 0.0590 \\
& & 0.05 & 0.1278 & 0.0560 & 0.0510 & 0.0532 & 0.0546 & 0.0691 & 0.1726 \\
& & 0.10 & 0.2080 & 0.1123 & 0.1069 & 0.1086 & 0.1051 & 0.1332 & 0.2650 \\
& {400} & 0.01 & 0.0324 & 0.0107 & 0.0099 & 0.0108 & 0.0097 & 0.0126 & 0.0467 \\
& & 0.05 & 0.1016 & 0.0500 & 0.0497 & 0.0560 & 0.0489 & 0.0592 & 0.1418 \\
& & 0.10 & 0.1734 & 0.1023 & 0.1002 & 0.1089 & 0.1017 & 0.1192 & 0.2232 \\
& {800} & 0.01 & 0.0252 & 0.0090 & 0.0098 & 0.0111 & 0.0102 & 0.0135 & 0.0354 \\
& & 0.05 & 0.0883 & 0.0504 & 0.0486 & 0.0478 & 0.0481 & 0.0618 & 0.1219 \\
& & 0.10 & 0.1555 & 0.1026 & 0.0986 & 0.0961 & 0.0997 & 0.1206 & 0.2044 \\ [1ex]
{$\Gamma(8,1)$} & {100} & 0.01 & 0.0488 & 0.0142 & 0.0117 & 0.0124 & 0.0111 & 0.0161 & 0.0576 \\
  & & .05 & 0.1425 & 0.0641 & 0.0611 & 0.0585 & 0.0589 & 0.0732 & 0.1687 \\
  & & 0.10 & 0.2241 & 0.1224 & 0.1147 & 0.1175 & 0.1150 & 0.1375 & 0.2679 \\
  & {200} & 0.01 & 0.0408 & 0.0100 & 0.0089 & 0.0122 & 0.0116 & 0.0143 & 0.0608 \\
  & & 0.05 & 0.1255 & 0.0559 & 0.0507 & 0.0534 & 0.0558 & 0.0645 & 0.1690 \\
  & & 0.10 & 0.1991 & 0.1105 & 0.1022 & 0.1048 & 0.1072 & 0.1233 & 0.2664 \\
  & {400} & 0.01 & 0.0314 & 0.0097 & 0.0114 & 0.0094 & 0.0103 & 0.0131 & 0.0430 \\
  & & 0.05 & 0.1002 & 0.0482 & 0.0521 & 0.0531 & 0.0561 & 0.0618 & 0.1404 \\
  & & 0.10 & 0.1697 & 0.1025 & 0.1066 & 0.1018 & 0.1100 & 0.1177 & 0.2232 \\
  & {800} & 0.01 & 0.0249 & 0.0092 & 0.0113 & 0.0112 & 0.0100 & 0.0113 & 0.0381 \\
  & & 0.05 & 0.0902 & 0.0488 & 0.0532 & 0.0511 & 0.0543 & 0.0598 & 0.1237 \\
  & & 0.10 & 0.1561 & 0.1024 & 0.1072 & 0.1034 & 0.1067 & 0.1177 & 0.1997 \\
   \bottomrule
\end{tabular}
\end{table}

@@

The KS test remains a crucial tool for researchers across diverse fields,
enabling critical assessments of population distributions. We introduced a
NPBB variant of the KS test tailored for data with
serial dependencies, which eliminates the need for explicit modeling serial
dependence in addition to parameter specification within the hypothesized
distribution. Our simulations have demonstrated that, given
sufficiently large samples of time series data and moderate serial dependencies,
the test maintains its size under the null hypothesis and have considerable
power in detecting deviations from the hypothesized marginal distribution. The
method's performance improves as the strength of temporal dependence decreases
or the sample size increases. For moderate serial dependence levels in practice,
the test can be recommended for sample size~$n = 400$ or higher.
Overall, the test provides an analog of the bias-corrected NPB KS test of
\citet{babu2004goodness} in the context of stationaly time
series with unspecified serial dependence. Compared to the SPB method
of \citet{zeimbekakis2022misuses}, which requires a working dependence model, it
is reliable with no need to worry about whether the working dependence
model approximates the true dependence model.

Several promising directions for further research emerge from the findings of
this study. Firstly, there are many other variants of bootstrap besides
circular block bootstrap which
can be used for serially dependent data such as the
stationary bootstrap \citep{politis1994stationary} and the dependent wild
bootstrap \citep{shao2010dependent}. It would be interesting to investigate how
these methods perform using our correction. Secondly, analogous
methodologies for other goodness-of-fit tests
based on empirical distributions, such as the Anderson--Darling and Cramér--von
Mises tests---which have been shown to outperform the KS test in certain scenarios
\citep{stephens2017tests}---could be explored. The application of block
bootstrapping bias correction to these tests could be particularly beneficial in
environmental studies, such as extreme value modeling, where observations are
often serially dependent. Thirdly, an extension of the proposed NPBB
one-sample KS test to a two-sample setting would be
intriguing. This adaptation would test whether two samples originate from the
same distribution without requiring specification of serial dependencies in
either sample, maintaining the advantageous feature of the original method.
Finally, a rigorous theoretical justification of the NPBB method for
stationary series is of significant academic interest. Establishing
precise conditions for the asymptotic behavior of the KS statistic and
the consistency of the bootstrap procedure would offer valuable
practical insights into the scenarios where the NPBB approach is
expected to perform optimally.

>>>>

%&latex
\documentclass[12pt]{article}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL
\usepackage{comment}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{setspace}

@@

\newpage

\doublespacing

\section{Introduction}\label{sec:intro}

@@

\begin{table}[tbp]
\centering
\caption{Empirical sizes or rejection rates
for the test that sample follows its true distribution family for
different values of AR(1) coefficient and for different
significance levels.}
\label{table:size}
\begin{tabular}{ccc ccc c ccc}
  \toprule
$F(\cdot; \theta)$ & $n$ & $\alpha$ & \multicolumn{7}{c}{Kendall's $\tau$}\\
\cmidrule(lr){4-10}
   &  &  & $-0.75$ & $ -0.50$ & $-0.25$ & $0.00$ & $0.25$ & $0.50$ & $ 0.75$ \\
  \midrule
{$N(8,8)$} & {100} & 0.01 & 0.0521 & 0.0124 & 0.0112 & 0.0106 & 0.0114 & 0.0158 & 0.0563 \\
& & 0.05 & 0.1470 & 0.0628 & 0.0545 & 0.0548 & 0.0575 & 0.0725 & 0.1705 \\
& & 0.10 & 0.2277 & 0.1186 & 0.1147 & 0.1072 & 0.1154 & 0.1424 & 0.2714 \\
& {200} & 0.01 & 0.0439 & 0.0113 & 0.0090 & 0.0114 & 0.0109 & 0.0138 & 0.0590 \\
& & 0.05 & 0.1278 & 0.0560 & 0.0510 & 0.0532 & 0.0546 & 0.0691 & 0.1726 \\
& & 0.10 & 0.2080 & 0.1123 & 0.1069 & 0.1086 & 0.1051 & 0.1332 & 0.2650 \\
& {400} & 0.01 & 0.0324 & 0.0107 & 0.0099 & 0.0108 & 0.0097 & 0.0126 & 0.0467 \\
& & 0.05 & 0.1016 & 0.0500 & 0.0497 & 0.0560 & 0.0489 & 0.0592 & 0.1418 \\
& & 0.10 & 0.1734 & 0.1023 & 0.1002 & 0.1089 & 0.1017 & 0.1192 & 0.2232 \\
& {800} & 0.01 & 0.0252 & 0.0090 & 0.0098 & 0.0111 & 0.0102 & 0.0135 & 0.0354 \\
& & 0.05 & 0.0883 & 0.0504 & 0.0486 & 0.0478 & 0.0481 & 0.0618 & 0.1219 \\
& & 0.10 & 0.1555 & 0.1026 & 0.0986 & 0.0961 & 0.0997 & 0.1206 & 0.2044 \\ [1ex]
{$\Gamma(8,1)$} & {100} & 0.01 & 0.0488 & 0.0142 & 0.0117 & 0.0124 & 0.0111 & 0.0161 & 0.0576 \\
  & & .05 & 0.1425 & 0.0641 & 0.0611 & 0.0585 & 0.0589 & 0.0732 & 0.1687 \\
  & & 0.10 & 0.2241 & 0.1224 & 0.1147 & 0.1175 & 0.1150 & 0.1375 & 0.2679 \\
  & {200} & 0.01 & 0.0408 & 0.0100 & 0.0089 & 0.0122 & 0.0116 & 0.0143 & 0.0608 \\
  & & 0.05 & 0.1255 & 0.0559 & 0.0507 & 0.0534 & 0.0558 & 0.0645 & 0.1690 \\
  & & 0.10 & 0.1991 & 0.1105 & 0.1022 & 0.1048 & 0.1072 & 0.1233 & 0.2664 \\
  & {400} & 0.01 & 0.0314 & 0.0097 & 0.0114 & 0.0094 & 0.0103 & 0.0131 & 0.0430 \\
  & & 0.05 & 0.1002 & 0.0482 & 0.0521 & 0.0531 & 0.0561 & 0.0618 & 0.1404 \\
  & & 0.10 & 0.1697 & 0.1025 & 0.1066 & 0.1018 & 0.1100 & 0.1177 & 0.2232 \\
  & {800} & 0.01 & 0.0249 & 0.0092 & 0.0113 & 0.0112 & 0.0100 & 0.0113 & 0.0381 \\
  & & 0.05 & 0.0902 & 0.0488 & 0.0532 & 0.0511 & 0.0543 & 0.0598 & 0.1237 \\
  & & 0.10 & 0.1561 & 0.1024 & 0.1072 & 0.1034 & 0.1067 & 0.1177 & 0.1997 \\
   \bottomrule
\end{tabular}
\end{table}

@@

The KS test remains a crucial tool for researchers across diverse fields,
enabling critical assessments of population distributions. We introduced a
NPBB variant of the KS test tailored for data with
serial dependencies, which eliminates the need for explicit modeling serial
dependence in addition to parameter specification within the hypothesized
distribution. Our simulations have demonstrated that, given
sufficiently large samples of time series data and moderate serial dependencies,
the test maintains its size under the null hypothesis and have considerable
power in detecting deviations from the hypothesized marginal distribution. The
method's performance improves as the strength of temporal dependence decreases
or the sample size increases. For moderate serial dependence levels in practice,
the test can be recommended for sample size~$n = 400$ or higher.
Overall, the test provides an analog of the bias-corrected NPB KS test of
\citet{babu2004goodness} in the context of stationary time
series with unspecified serial dependence. Compared to the SPB method
of \citet{zeimbekakis2022misuses}, which requires a working dependence model, it
is reliable with no need to worry about whether the working dependence
model approximates the true dependence model.

Several promising directions for further research emerge from
this study. Firstly, while we used circular block bootstrap, many
other bootstrap variants for serially dependent data, such as the
stationary bootstrap \citep{politis1994stationary} and the dependent wild
bootstrap \citep{shao2010dependent}, could be evaluated under our
correction framework. Secondly, analogous methodologies could be
developed for other goodness-of-fit tests based on empirical
distributions, including the Anderson---Darling and Cramér--von
Mises tests, which are known to outperform the KS test in certain scenarios
\citep{stephens2017tests}. Applying block bootstrap bias correction to
these tests may be particularly beneficial in contexts like extreme
value modeling, where serial dependence is common. Third, it would be
intriguing to extend the proposed NPBB
one-sample KS test to a two-sample setting, allowing for comparison of
marginal distributions without specifying serial dependencies in
either sample. Finally, a rigorous theoretical justification of the
NPBB procedure, including precise conditions for the asymptotic
behavior of the test statistic and the consistency of the bootstrap
approximation, would offer valuable practical insights into the
scenarios where the NPBB approach is expected to perform optimally.
